---
title: "class09"
author: "Nguyen"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

The main k-means function in R is called `kmeans()`

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
print (x)
plot(x)
```
Use the kmeans() function setting k to 2 and nstart=20
 
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```
 
Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?

```{r}
km$size

km$cluster
length(km$cluster)
table(km$cluster)
```
 
 Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
plot(x, col=km$cluster)
points(km$centers, col="Blue", pch = 15, cex = 3)
```
 
 
 
##Hierarchical clustering in R

The main Hierarchical clustering function in R is called `hclust()`
An important point here is that you have to calculate the distance matric deom your input data before calling `hclust`
 
```{r}
#we will use our x again from above
d <- dist(x)
hc <- hclust(d)
hc
plot(hc)
abline(h=6, col="red", lty=2)
abline(h=4, col="blue")

```
 
 To get cluster membership vector, I need to "cut" the tree at a certain height to yield my seperate cluster branches.
 
```{r}
cutree(hc, h=6)
```
 
```{r}
cutree(hc, h=4)
```
 
```{r}
cutree(hc, k=2)
```
 
 
 
 
## Linking clusters

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
Q. How does this compare to your known 'col' groups?
```{r}
hc <- hclust(dist(x))
plot(hc)
abline(h = 1.7, col = "red")
```

To get cluster membership vector use 'cutree'

```{r}
grps <- cutree(hc, k=3)
table(grps)
plot(x, col=grps)
```




#Hands on class09
```{r}
x <- read.csv("UK_foods.csv", row.names=1)
head(x)
dim(x)
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

Principal Component Analysis (PCA) with the `prcomp()` function

```{r}
pca <- prcomp( t(x))
```



what is my result object `pca`? I can check the attributes...

```{r}
attributes(pca)
```

```{r}
plot(pca$x[,1],pca$x[,2])
```

```{r}
summary(pca)
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```


