---
title: "class12"
author: "Nguyen"
date: "2/13/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PDB statistics

Here we inspect the types of structures in the main database for 3D biomolecular data - the PDB

Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. 

>Q1. Determine the percentage of structures solved by X-Ray and Electron Microscopy.

```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

#calculate percetnage values for each method

```


>Q2. Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

>Q2  Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?