---
title: "class10"
author: "Nguyen"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of Human Breast Cancer Cells

First we need to important (i.e. read) our input data. I downloaded the csv file from the class website:
https://bioboot.github.io/bimm143_W20/class-material/WisconsinCancer.csv

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```
There are some funky things in this dataset that we will ignore for our analysis. This includes the first and second ID and Diagnosis columns and the funny last X column (col 35)

```{r}
#Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)
```

>Q. How many patients do we have data for?

```{r}
nrow(wisc.data)
```

>Q. How many cancer and non-cancer?

```{r}
table(wisc.df$diagnosis)
```

>Q. How many variables/features in the data are suffixed with _mean

For this I will turn to the `grep()` function and look at the help page to see how it works.

```{r}
grep("_mean", colnames(wisc.data), value=TRUE)
```


```{r}
grep("_mean", colnames(wisc.data), value=TRUE)
```

##Principal Component Analysis

Before we turn to PCA we need to think, or consider, whether we should SCALE our input.








```{r}
round(apply(wisc.data, 2,sd), 2)
```


Looks like we need to set scale = TRUE

```{r}
#Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

The 1st PC captures 44.27% of the original variance (look at cumulative proportion). Note that 72.6% is captured in the first 3 PCs.



Lets make some figures..

```{r}
biplot(wisc.pr)
```

That is a hot mess! We need to do out own PC1 vs PC2 plot and lets color by the expert diagnosis.

```{r}
attributes(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)
```

Cluster in PC space

First, lets see if we can cluster the original data

```{r}
wisc.hc <- hclust(dist(wisc.data))
plot(wisc.hc)
```







```{r}
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:3]), method="ward.D2")
plot(wisc.pr.hclust)
```

To get our clusters out of this tree we need to CUT it with `cutree()` function.

```{r}
grps3 <- cutree(wisc.pr.hclust, k=3)
table(grps3)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$pr$x[,2], col=grps3)
```

We can use the `table()` function to compare the $diagnosis vector with our cluster results vector.

```{r}
table(grps3, wisc.df$diagnosis)
```




##Predcition

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Now draw the PCA plot again and add our new data:

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col="white")
```

