---
title: "class 15"
author: "Nguyen"
date: "2/20/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

How many genes are we dealing with?

```{r}
nrow(counts)
```

We want to calculate the average counts per gene for control samples and for the drug treated samples. We will compare these values to see if there is a difference in gene expression.

Now I want access to the control only column in the metadata
```{r}
control <- metadata[metadata$dex == "control",]
control
```


Now use the $id of the control data to access the control collumns in `counts` and calculate the average.
```{r}
counts[, control$id]
```

```{r}
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```

Now lets do the same for the *treated* samples

```{r}
treated <- metadata[metadata$dex == "treated",]
treated
```

```{r}
counts[, treated$id]
```
```{r}
treated.mean <- rowSums(counts[ ,treated$id] )/4 
names(treated.mean) <- counts$ensgene
```


We will combine our meancount data for bookkeeping purposes.
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

```{r}
plot(meancounts)
```


#fold change
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```


Remove any gene that has zero average counts.

```{r}
x <- data.frame(c(1,1,10,0), c(0,1,10,0))
x
x==0
ind <- unique(which(x == 0, arr.ind = TRUE)[,"row"])
x[-ind]
```


How many genes do I have left?

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes do I have left
```{r}
nrow(mycounts)
```

A common threshold used for calling something differently expressed is a log2(FoldChange)
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```


```{r}
sum(up.ind)
sum(down.ind)
```

## Do this properly with DESeq2

```{r}
library(DESeq2)
```


Setup input for DESeq by specifying countdata, metadata, and the experiment design (i.e. which comparisons we are interested in)

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
res <- results(dds)
```

```{r}
plot(res$log2FoldChange, res$padj)
```

Lets plot this the way the rest of the world does!
```{r}
plot( res$log2FoldChange,  -log(res$padj))
```

# Set the color palette for our plot
```{r}
mycols <- rep("gray", nrow(res))
mycols[ res$log2FoldChange > 2 ] <- "blue"
mycols[ res$log2FoldChange < -2 ] <- "blue"
plot(res$log2FoldChange, -log(res$padj), col=mycols)
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)
```






