---
title: 'Class 6: R Functions'
author: "Nguyen"
date: "1/23/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Level 2 heading

### Level 3 heading

```{r}
# this is a silly graph
plot(1:10, col = "lightblue")
```

Let's see more about **file import** (i.e. reading files into R). The main read function in base R is 'read.table()'

```{r}
t1 <- read.table("test1.txt", header = TRUE, sep = ",")
```

Or I could just use 'read.csv()' which has the arguements I want in this case!
```{r}
t1 <- read.csv("test1.txt")

t2 <- read.table("test2.txt", sep = "$")

t3 <- read.table("test3.txt")
```

#Back to functions

Our first example function

```{r}
add <- function(x, y=1){
  #Sum the input x and y
  x + y
}
```

```{r}
add(7,3)
```

#How does this work with vectors as input.
```{r}
add( c(1,2,3),4 )
```
```{r}
add( c(1,2,3), c(1,2,3) )
```


What is this `range()` function you speak of?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)

range(x)
```

This is our second function:
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(x)
```

Do some more testing
```{r}
rescale(1:10)
```
```{r}
x <- c(1,2,NA,3,10)

rescale2 <- function(x, na.rm=TRUE) {
 rng <-range(x, na.rm = na.rm)
 (x - rng[1]) / (rng[2] - rng[1])
}

rescale2(x)

rescale2(x, na.rm =FALSE)
```


```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```

```{r}
rescale3(x)
```


Hands On Section 1:B
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```


Hands On Questions


```{r}
read.pdb("4AKE") # kinase with drug
```

>Q1. What ype of object is returned from `read.pdb
It is a list of eight things

```{r}
help(read.pdb)
```


>Q4

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, type = "l", col = "blue", lwd=2)
points(s3.b, type = "l", col = "red", lwd=2)
```



>Q3:What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?
# Can you improve this analysis code?

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, typ="l", ylab="Bfactor")
plotb3(s2.b, typ="l", ylab="Bfactor")
plotb3(s3.b, typ="l", ylab="Bfactor")
```


>Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) ) 
plot(hc)
```

```{r}
library(bio3d)
files <- c("4AKE", "1AKE", "1E4Y")
for (i in 1:length(files)) {
s1 <- read.pdb(files[i])
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")}
```



